---
title: "Mapping and Measuring Residential Segregation"
author: "Arpita Banerjee."
date: "3/19/2021"
output:
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  include: FALSE
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE, message=FALSE}
rm(list=ls())
#census_api_key("936c96236b979ae522c6cf67edb51923cd391fb3")

knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(tidycensus)
library(sf)
library(tigris)


#Downloading census data
B01001_Vars <- c("B01001_001")
B01001 <-
  get_acs(
    geography = "tract",
    state = "Illinois",
    variables = B01001_Vars,
    survey = "acs5",
    year = 2019,
    output = "wide"
  )
B01001$Pop <- B01001$B01001_001E
B01001$Pop[B01001$Pop == "NaN"] <- NA
B01001 <-
  B01001 %>% filter(!is.na(Pop), na.rm = TRUE) %>% filter(Pop > 0) %>% select(GEOID, Pop)

#Race
B02001_Vars <- c(
  "B02001_001",
  "B02001_002",
  "B02001_003",
  "B02001_004",
  "B02001_005",
  "B02001_006",
  "B02001_007",
  "B02001_008"
)
B02001 <-
  get_acs(
    geography = "tract",
    state = "Illinois",
    variables = B02001_Vars,
    year = 2019,
    survey = "acs5",
    output = "wide"
  )
B02001$White <- B02001$B02001_002E
B02001$Black <- B02001$B02001_003E
B02001$AIAN <- B02001$B02001_004E
B02001$Asian <- B02001$B02001_005E
B02001$Nonwhite <- (B02001$B02001_001E - B02001$B02001_002E)
B02001$PWhite <- B02001$White / B02001$B02001_001E
B02001$PBlack <- B02001$Black / B02001$B02001_001E
B02001$PAIAN <- B02001$AIAN / B02001$B02001_001E
B02001$PAsian <- B02001$Asian / B02001$B02001_001E
B02001$PNonwhite <- B02001$Nonwhite / B02001$B02001_001E
B02001$PWhite[B02001$PWhite == "NaN"] <- NA
B02001$PBlack[B02001$PBlack == "NaN"] <- NA
B02001$PAIAN[B02001$PAIAN == "NaN"] <- NA
B02001$PAsian[B02001$Asian == "NaN"] <- NA
B02001$PNonwhite[B02001$Nonwhite == "NaN"] <- NA
B02001 <-
  B02001 %>% select(GEOID,
                    White,
                    Black,
                    AIAN,
                    Asian,
                    Nonwhite,
                    PWhite,
                    PBlack,
                    PAIAN,
                    PAsian,
                    PNonwhite)

#Ethnicity 
B03001_Vars <- c("B03001_001", "B03001_003")
B03001 <-
  get_acs(
    geography = "tract",
    state = "Illinois",
    variables = B03001_Vars,
    year = 2019,
    survey = "acs5",
    output = "wide"
  )
B03001$Latino <- B03001$B03001_003E
B03001$PLatino <- B03001$Latino / B03001$B03001_001E
B03001$PLatino[B03001$PLatino == "NaN"] <- NA
B03001 <- B03001 %>% select(GEOID, Latino, PLatino)

#B19013: Median Household Income 
B19013_Vars <- c("B19013_001")
B190193 <-
  get_acs(
    geography = "tract",
    state = "Illinois",
    variables = B19013_Vars,
    survey = "acs5",
    year = 2019,
    output = "wide"
  )
B190193$MHHI <- B190193$B19013_001E
B190193$MHHI[B190193$MHHI == "NaN"] <- NA
B190193 <- B190193 %>% select(GEOID, MHHI)

# Join the data together and clean up intermediate data objects.
acs_data <- left_join(B01001, B02001, by = "GEOID")
acs_data <- left_join(acs_data, B03001, by = "GEOID")
acs_data <- left_join(acs_data, B190193, by = "GEOID")
rm(B01001,B02001, B03001, B190193, B01001_Vars, B02001_Vars, B03001_Vars, B19013_Vars)
```

# 1. Introduction 

The data used in the following analysis has been taken from the US Census Bureau, American Community Survey for 5 year periods. It is based on the tables described below. All the tables have been directly imported into R through Tidycensus – 
| ACS Table | Variables                | Label    | Description                                            |
|-----------|--------------------------|----------|--------------------------------------------------------|
| B01001    | B01001_001E              | Pop      | Total Population                                       |
| B02001    | B02001_002E              | White    | Count of White Population                              |
| B02001    | B02001_003E              | Black    | Count of Black Population                              |
| B02001    | B02001_004E              | AIAN     | Count of American Indian and Alaskan Native Population |
| B02001    | B02001_005E              | Asian    | Count of Asian Population                              |
| B02001    | B02001_001E, B02001_002E | Nonwhite | Count of Nonwhite Population                           |
| B03001    | B03001_003E              | Latino   | Count of Latino Population                             |
| B19013    | B19013_001E              | MHHI     | Median Household Income                                |

```{r, echo=FALSE, message=FALSE}
#Downloading tract level spatial data
il_trt <- tracts(state="IL", class="sf")
ggplot() + geom_sf(data = il_trt)+labs(title = "Tracts in Illinois, 2019")
il_trt <- left_join(il_trt, acs_data, by="GEOID")
il_trt <- il_trt %>% filter(!is.na(Pop), na.rm=TRUE)
```

## 1.1. Chicago Metropolitan Region

The primary aim of this analysis is to understand how households are segregated in the Chicago Metropolitan Region, spanning the following counties
| County  | FIPS |
|---------|------|
| Cook    | 031  |
| DeKalb  | 037  |
| DuPage  | 043  |
| Grundy  | 063  |
| Kane    | 089  |
| Kendall | 093  |
| Lake    | 097  |
| McHenry | 111  |
| Will    | 197  |
Several tracts within the aforementioned counties are uninhabited and are therefore removed from the analysis of housing segregation described below. 
```{r, echo=FALSE}
chi <- il_trt %>% filter(COUNTYFP %in% c("031", "037", "043", "063", "089", "093", "097", "111", "197"))
ggplot() + geom_sf(data = chi) + labs(title = "Tracts [with Population] in Chicago Metropolitan Region, 2019")
```
# 2. Distribution of Racial Groups
```{r, echo=FALSE}
co <- chi %>% group_by(COUNTYFP) %>% 
  summarise(Pop_co=sum(Pop),
            White_co=sum(White),
            Black_co=sum(Black),
            AIAN_co=sum(AIAN),
            Asian_co=sum(Asian),
            Latino_co=sum(Latino),
            Nonwhite_co=sum(Nonwhite))

ggplot() + 
  geom_sf(data = chi, aes(fill = PWhite)) + scale_fill_gradient(low = "0.00", high = "1.00") + 
  geom_sf(data = co,
          colour = "RED",
          fill = NA) + labs(title = "% of Whites in Chicago Metropolitan Region by County, 2019")
```

## 2.1. Mapping Racial Minorities In Chicago Metropolitan Region
```{r, echo=FALSE}
ggplot() + 
  geom_sf(data = chi, aes(fill = PNonwhite)) + scale_fill_gradient(low = "0.00", high = "1.00") + 
  geom_sf(data = co,
          colour = "RED",
          fill = NA) + labs(title = "% of Non Whites in Chicago Metropolitan Region by County, 2019")

```

```{r, echo=FALSE}
ggplot() + 
  geom_sf(data = chi, aes(fill = PBlack)) + scale_fill_gradient(low = "0.00", high = "1.00") + 
  geom_sf(data = co,
          colour = "RED",
          fill = NA) + labs(title = "% of Blacks in Chicago Metropolitan Region by County, 2019")

```

```{r, echo=FALSE}
ggplot() + 
  geom_sf(data = chi, aes(fill = PLatino)) + scale_fill_gradient(low = "0.00", high = "1.00") + 
  geom_sf(data = co,
          colour = "RED",
          fill = NA) + labs(title = "% of Latinos in Chicago Metropolitan Region by County, 2019")

```

```{r, echo=FALSE}
ggplot() + 
  geom_sf(data = chi, aes(fill = PAsian)) + scale_fill_gradient(low = "0.00", high = "1.00") + 
  geom_sf(data = co,
          colour = "RED",
          fill = NA) + labs(title = "% of Asians in Chicago Metropolitan Region by County, 2019")

```

```{r, echo=FALSE}
ggplot() + 
  geom_sf(data = chi, aes(fill = PAIAN)) + scale_fill_gradient(low = "0.00", high = "1.00") + 
  geom_sf(data = co,
          colour = "RED",
          fill = NA) + labs(title = "% of American Indians in Chicago Metropolitan Region by County, 2019")

```
## 2.2. Comparing Relative Distribution of Different Racial Groups
```{r, echo=FALSE}
co %>% ggplot(aes(y=White_co, x=COUNTYFP))+geom_col()+theme_minimal()+ geom_text(aes(label = White_co), vjust = -0.5)+labs(title = "White Population at the County Level, Chicago Metropolitan Region, 2019")
```

```{r, echo=FALSE}
co %>% ggplot(aes(y=Nonwhite_co, x=COUNTYFP))+geom_col()+theme_minimal()+ geom_text(aes(label = Nonwhite_co), vjust = -0.5)+labs(title = "Non White Population at the County Level, Chicago Metropolitan Region, 2019")
```

```{r, echo=FALSE}
co %>% ggplot(aes(y=Black_co, x=COUNTYFP))+geom_col()+theme_minimal()+ geom_text(aes(label = Black_co), vjust = -0.5)+labs(title = "Black Population at the County Level, Chicago Metropolitan Region, 2019")
```

```{r, echo=FALSE}
co %>% ggplot(aes(y=Latino_co, x=COUNTYFP))+geom_col()+theme_minimal()+ geom_text(aes(label = Latino_co), vjust = -0.5)+labs(title = "Latino at the County Level, Chicago Metropolitan Region, 2019")
```

```{r, echo=FALSE}
co %>% ggplot(aes(y=Asian_co, x=COUNTYFP))+geom_col()+theme_minimal()+ geom_text(aes(label = Asian_co), vjust = -0.5)+labs(title = "Asian Population at the County Level, Chicago Metropolitan Region, 2019")
```

```{r, echo=FALSE}
co %>% ggplot(aes(y=AIAN_co, x=COUNTYFP))+geom_col()+theme_minimal()+ geom_text(aes(label = AIAN_co), vjust = -0.5)+labs(title = "American Indian Population at the County Level, Chicago Metropolitan Region, 2019")
```
# 3. Understanding Housing Segregation
```{r, echo=FALSE}

```
## 3.1. Majority/Minority Dissimilarity
```{r, echo=FALSE, message=FALSE}
chi <- left_join(chi, co %>% st_set_geometry(NULL), by = "COUNTYFP")

dissimilarity_major_minor_NonWhite <- chi %>% mutate(dissim_nww = abs(Nonwhite / Nonwhite_co - White / White_co)) %>%
  group_by(COUNTYFP) %>%
  summarise(dissimilarity_NonWhite_White = 0.5 * sum(dissim_nww))%>% st_set_geometry(NULL)
dissimilarity_major_minor_Black <- chi %>% mutate(dissim_wb = abs(Black / Black_co - White / White_co)) %>%
  group_by(COUNTYFP) %>%
  summarise(dissimilarity_Black_White = 0.5 * sum(dissim_wb))%>% st_set_geometry(NULL)
dissimilarity_major_minor_Latino <- chi %>% mutate(dissim_lw = abs(Latino / Latino_co - White / White_co)) %>%
  group_by(COUNTYFP) %>%
  summarise(dissimilarity_Latino_White = 0.5 * sum(dissim_lw))%>% st_set_geometry(NULL)
dissimilarity_major_minor_Asian<- chi %>% mutate(dissim_Asian = abs(Asian / Asian_co - White / White_co)) %>%
  group_by(COUNTYFP) %>%
  summarise(dissimilarity_Asian_White = 0.5 * sum(dissim_Asian))%>% st_set_geometry(NULL)
dissimilarity_major_minor_AIAN<- chi %>% mutate(dissim_AIAN = abs(AIAN / AIAN_co - White / White_co)) %>%
  group_by(COUNTYFP) %>%
  summarise(dissimilarity_AIAN_White = 0.5 * sum(dissim_AIAN))%>% st_set_geometry(NULL)

dissimilarity_major_minor <- left_join(dissimilarity_major_minor_NonWhite, dissimilarity_major_minor_Black, by = "COUNTYFP")
dissimilarity_major_minor <- left_join(dissimilarity_major_minor, dissimilarity_major_minor_Latino, by = "COUNTYFP")
dissimilarity_major_minor <- left_join(dissimilarity_major_minor, dissimilarity_major_minor_Asian, by = "COUNTYFP")
dissimilarity_major_minor <- left_join(dissimilarity_major_minor, dissimilarity_major_minor_AIAN, by = "COUNTYFP")

rm(dissimilarity_major_minor_AIAN, dissimilarity_major_minor_Asian, dissimilarity_major_minor_Black, dissimilarity_major_minor_Latino, dissimilarity_major_minor_NonWhite)
knitr::kable(dissimilarity_major_minor)
```
## 3.2. Interaction Between Different Racial Groups
```{r, echo=FALSE, message=FALSE}
interact_NonWhite_white <- chi %>% mutate(interact_nnw = abs(Nonwhite / Nonwhite_co * White / Pop)) %>%
  group_by(COUNTYFP) %>%
  summarise(interaction_NonWhite_White = sum(interact_nnw))%>% st_set_geometry(NULL)
interact_black_white <- chi %>% mutate(interact_bw = abs(Black / Black_co * White / Pop)) %>%
  group_by(COUNTYFP) %>%
  summarise(interaction_Black_White = sum(interact_bw))%>% st_set_geometry(NULL)
interact_Latino_white <- chi %>% mutate(interact_lw = abs(Latino / Latino_co * White / Pop)) %>%
  group_by(COUNTYFP) %>%
  summarise(interaction_Latino_White = sum(interact_lw))%>% st_set_geometry(NULL)
interact_Asian_white <- chi %>% mutate(interact_aw = abs(Asian / Asian_co * White / Pop)) %>%
  group_by(COUNTYFP) %>%
  summarise(interaction_Asian_White = sum(interact_aw))%>% st_set_geometry(NULL)
interact_AIAN_white <- chi %>% mutate(interact_aiw = abs(AIAN / AIAN_co * White / Pop)) %>%
  group_by(COUNTYFP) %>%
  summarise(interaction_Asian_White = sum(interact_aiw))%>% st_set_geometry(NULL)
interaction_overall <- left_join(interact_NonWhite_white, interact_black_white, by = "COUNTYFP")
interaction_overall <- left_join(interaction_overall, interact_Latino_white, by = "COUNTYFP")
interaction_overall <- left_join(interaction_overall, interact_Asian_white, by = "COUNTYFP")
interaction_overall <- left_join(interaction_overall, interact_AIAN_white, by = "COUNTYFP")

rm(interact_NonWhite_white, interact_black_white, interact_Latino_white, interact_Asian_white, interact_AIAN_white)
knitr::kable(interaction_overall)
```
## 3.3. Isolation of Minorities
```{r, echo=FALSE, message=FALSE}
isolation_nw <- chi %>% mutate(iso_nw = abs(Nonwhite / Nonwhite_co * Nonwhite / Pop)) %>%
  group_by(COUNTYFP) %>% summarise(NonWhite_Isolation = sum(iso_nw)) %>% st_set_geometry(NULL)
isolation_b <- chi %>% mutate(iso_b = abs(Black / Black_co * Black / Pop)) %>%
  group_by(COUNTYFP) %>% summarise(Black_Isolation = sum(iso_b)) %>% st_set_geometry(NULL)
isolation_l <- chi %>% mutate(iso_l = abs(Latino / Latino_co * Latino / Pop)) %>%
  group_by(COUNTYFP) %>% summarise(Latino_Isolation = sum(iso_l)) %>% st_set_geometry(NULL)
isolation_a <- chi %>% mutate(iso_a = abs(Asian / Asian_co * Asian / Pop)) %>%
  group_by(COUNTYFP) %>% summarise(Asian_Isolation = sum(iso_a)) %>% st_set_geometry(NULL)
isolation_ai <- chi %>% mutate(iso_ai = abs(AIAN / AIAN_co * AIAN / Pop)) %>%
  group_by(COUNTYFP) %>% summarise(AIAN_Isolation = sum(iso_ai)) %>% st_set_geometry(NULL)

isolation_all <- left_join(isolation_nw, isolation_b, by = "COUNTYFP")
isolation_all <- left_join(isolation_all, isolation_l, by = "COUNTYFP")
isolation_all <- left_join(isolation_all, isolation_a, by = "COUNTYFP")
isolation_all <- left_join(isolation_all, isolation_ai, by = "COUNTYFP")

rm(isolation_nw, isolation_b, isolation_l, isolation_a, isolation_ai)
knitr::kable(isolation_all)
```
# 4. Intersectionality between Forms of Segregation - Income and Race
```{r, echo=FALSE}

```
## 4.1. Income Segregation
```{r, echo=FALSE}
rm(county_median_income)
county_var <- c("S1901_C01_012")
county_median_income <-
  get_acs(
    geography = "county",
    state = "Illinois",
    year = 2019,
    variables = county_var,
    survey = "acs5",
    output = "wide"
  )
county_median_income <-
  county_median_income %>% filter(
    NAME %in% c(
      "Cook County, Illinois",
      "DeKalb County, Illinois",
      "DuPage County, Illinois",
      "Grundy County, Illinois",
      "Kane County, Illinois",
      "Kendall County, Illinois",
      "Lake County, Illinois",
      "McHenry County, Illinois",
      "Will County, Illinois"
    )
  )
county_median_income <-
  county_median_income %>% mutate(
    COUNTYFP = case_when(
      NAME %in% c("Cook County, Illinois") ~ "031",
      NAME == "DeKalb County, Illinois" ~ "037",
      NAME == "DuPage County, Illinois" ~ "043",
      NAME == "Grundy County, Illinois" ~ "063",
      NAME == "Kane County, Illinois" ~ "089",
      NAME == "Kendall County, Illinois" ~ "093",
      NAME == "Lake County, Illinois" ~ "097",
      NAME == "McHenry County, Illinois" ~ "111",
      NAME == "Will County, Illinois" ~ "197"
    )
  )

county_median_income <- county_median_income %>% rename(Median_Household_Income = S1901_C01_012E) %>% select(GEOID, NAME,Median_Household_Income,COUNTYFP )

chi <- left_join(chi,county_median_income, by="COUNTYFP") 
chi <- chi %>% mutate(inc_ratio = MHHI/Median_Household_Income)
ggplot() + geom_sf(data = chi, aes(fill = inc_ratio))
```

## 4.2. Income and Racial Segregation
```{r, echo=FALSE}
ggplot() + geom_point(data = chi, aes(x = PWhite, y = inc_ratio)) + geom_abline(method = "lm") +
  labs(title = "Population of Whites v/s Median Household Income Ratios, 2019", x = "Percent White", y = "Income Ratio")
```

```{r, echo=FALSE}
ggplot() + geom_point(data = chi, aes(x = PWhite, y = inc_ratio)) + geom_abline(method = "lm") +
  labs(title = "Whites v/s Income Ratios for all Chicago Metropolitan region Counties, 2019", x = "Percent White", y = "Income Ratio") +
  facet_wrap(~ NAME.y, scales = "free") + theme_minimal()

```

```{r, echo=FALSE}
dissimilarity_major_minor_NonWhite <- chi %>% mutate(dissim_nww = abs(Nonwhite / Nonwhite_co - White / White_co)) %>%
  group_by(COUNTYFP) %>%
  summarise(dissimilarity_NonWhite_White = 0.5 * sum(dissim_nww))%>% st_set_geometry(NULL)
interact_NonWhite_white <- chi %>% mutate(interact_nnw = abs(Nonwhite / Nonwhite_co * White / Pop)) %>%
  group_by(COUNTYFP) %>%
  summarise(interaction_NonWhite_White = sum(interact_nnw))%>% st_set_geometry(NULL)
isolation_NonWhite <- chi %>% mutate(iso_nw = abs(Nonwhite / Nonwhite_co * Nonwhite / Pop)) %>%
  group_by(COUNTYFP) %>% summarise(NonWhite_Isolation = sum(iso_nw)) %>% st_set_geometry(NULL)


income_segregation <- left_join(dissimilarity_major_minor_NonWhite, interact_NonWhite_white, by= "COUNTYFP")
income_segregation <- left_join(income_segregation,isolation_NonWhite, by= "COUNTYFP")
income_segregation <- left_join(income_segregation, county_median_income,by = "COUNTYFP" )
income_segregation <- income_segregation %>% select(COUNTYFP, NAME, Median_Household_Income, dissimilarity_NonWhite_White, interaction_NonWhite_White, NonWhite_Isolation)

knitr::kable(income_segregation)
```



## 4.3. Other Forms of Segregation

# 5. Conclusions

